|

### 2020-10-30 pre-analysis-plan/freire-skarbek-lynching.Rmd
add residency category, small edits

[^thanks]: We thank Rosario Aguilar, Louise Araújo, Miriam Golden, Umberto
Mignozzetti, Irfan Nooruddin, Catarina Roman, and Alex Scacco for their valuable comments.

@@

In this study, we endeavour to answer the following questions: 1) which victim
characteristics increase individual support for extralegal punishment?; 2) how
do individuals justify their approval for vigilantism?; and 3) can information
provision reduce support for mob violence? We design a set of three experiments
to tackle these issues. Our first experiment consists of a conjoint analysis
in which we show respondents different profiles of lynching victims and ask them
who they believe deserve punishment. In our second experiment, we test how
ineffective policing, slow criminal justice systems, and demand for hasher legal
punishment affect individual propensity to support mob violence. Finally, we
assess whether citizens become less supportive of vigilantism when they are
informed about the legal consequences of lynchings, constitutional human rights,
or the risk of _vendettas_ associated with mob violence. While in this
pre-analysis plan we present the three experiments as described above,
their order will be randomised in our survey to mitigate potential order bias.

@@

We start our analysis with a choice-based conjoint experimental design. We will
present five pairs of criminal profiles to respondents. Each profile consists
of seven attributes: 1) gender of the crime perpetrator; 2) age of the crime
perpetrator; 3) race of the crime perpetrator; 4) offense; 5) gender of the victim
of the motivating crime; 6) age of the victim of the motivating crime; 7)
lynching perpetrators. The attributes and levels are displayed in table
\ref{tab:categories} below.

@@

\begin{table}[htpb]
\small
\begin{center}
\caption{\textbf{Attributes and Levels}}
\label{tab:categories}
\begin{tabular}{l !{\vrule width 1pt}p{9cm}}
\Xhline{2\arrayrulewidth}
\textbf{Attribute} & \multicolumn{1}{c}{\textbf{Levels}} \\
\Xhline{2\arrayrulewidth}
\small
Gender of crime perpetrator & Male; female \\ [4pt]
Age of crime perpetrator & Teenager; adult; elderly \\ [4pt]
Race of crime perpetrator & Black; White; Native Brazilian; Asian \\ [4pt]
Offense & Picks the pocket; steals the car; molests; rapes; murders \\ [4pt]
Gender of crime victim & Male; female\\ [4pt]
Age of crime victim & Child; teenager; adult; elderly\\ [4pt]
Lynching perpetrators & Bystanders; neighbours; family of the victim; gangs; police \\
\Xhline{2\arrayrulewidth}
\end{tabular}
\end{center}
\end{table}

@@

Respondents have to indicate which profile they prefer for extrajudicial
punishment. They are are also required to rate each profile on a scale from 1 to
7, where 1 indicates they definitely believe lynching was not justified and 7
indicates they definitely believe lynching was justified. We take this measure
as a robustness check for our main experimental estimate.

@@

1) Gender of crime perpetrator: male 
2) Age of crime perpetrator: teenager
3) Race of crime perpetrator: blacks
4) Offense: murder
5) Gender of victim: female 
6) Age of victim: child
7) Lynching perpetrators: the family of the victim

@@

* *Treatment 02 - Criminal justice ineffectiveness*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial system is too slow"**.[^treat02-exp02]

[^treat02-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a justiça
é muito lenta"**.

@@

* You will be shown a news article. Please read it carefully. After you read
  the article, we will ask one question about it.[^vig01-exp02]

>>>>

[^thanks]: We thank Rosario Aguilar, Louise Araújo, Miriam Golden, Umberto
Mignozzetti, Irfan Nooruddin, Catarina Roman, and Alex Scacco for their
valuable comments. This pre-analysis plan received IRB approval from Brown
University in October 2020 (Protocol 2009002803).

@@

In this study, we endeavour to answer the following questions: 1) which victim
characteristics increase individual support for extralegal punishment?; 2) how
do individuals justify their approval for vigilantism?; and 3) can information
provision reduce support for mob violence? We design a set of three experiments
to tackle these issues. Our first experiment consists of a conjoint analysis in
which we show respondents different profiles of lynching victims and ask them
who they believe deserve punishment. In our second experiment, we test how
ineffective policing, slow criminal justice systems, and demand for hasher
legal punishment affect individual propensity to support mob violence. Finally,
we assess whether citizens become less supportive of vigilantism when they are
informed about the legal consequences of lynchings, constitutional human
rights, or the risk of _vendettas_ associated with mob violence. While in this
pre-analysis plan we present the three experiments as described above, we will
randomised their order in our survey to mitigate potential order bias.

@@

We start our analysis with a choice-based conjoint experimental design. We will
present five pairs of criminal profiles to respondents. Each profile consists
of eight attributes: 1) gender of the crime perpetrator; 2) age of the crime
perpetrator; 3) race of the crime perpetrator; 4) residency of crime
perpetrator; 5) offense; 6) gender of the victim of the motivating crime; 7)
age of the victim of the motivating crime; 8) lynching perpetrators. The
attributes and levels are displayed in table \ref{tab:categories} below.

@@

\begin{table}[htpb]
\small
\begin{center}
\caption{\textbf{Attributes and Levels}}
\label{tab:categories}
\begin{tabular}{l !{\vrule width 1pt}p{9cm}}
\Xhline{2\arrayrulewidth}
\textbf{Attribute} & \multicolumn{1}{c}{\textbf{Levels}} \\
\Xhline{2\arrayrulewidth}
\small
Gender of crime perpetrator & Male; female \\ [4pt]
Age of crime perpetrator & Teenager; adult; elderly \\ [4pt]
Race of crime perpetrator & Black; White; Native Brazilian; Asian \\ [4pt]
Residency of crime perpetrator & Resident in the community; outsider \\ [4pt]
Offense & Picks the pocket; steals the car; molests; rapes; murders \\ [4pt]
Gender of crime victim & Male; female\\ [4pt]
Age of crime victim & Child; teenager; adult; elderly\\ [4pt]
Lynching perpetrators & Bystanders; neighbours; family of the victim; gangs; police \\
\Xhline{2\arrayrulewidth}
\end{tabular}
\end{center}
\end{table}

@@

Respondents have to indicate which profile they prefer for extrajudicial
punishment. We will also include a qualitative textbox below each experiment
for subjects to explain why they selected their case or whether they believe
neither alternative is appropriate.

@@

1) Gender of crime perpetrator: male 
2) Age of crime perpetrator: teenager
3) Race of crime perpetrator: blacks
4) Residency of crime perpetrator: outsider
5) Offense: murder
6) Gender of victim: female 
7) Age of victim: child
8) Lynching perpetrators: the family of the victim

@@

* *Treatment 02 - Criminal justice ineffectiveness*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial system is too slow and the perpetrator is on
  the street until the case is heard"**.[^treat02-exp02]

[^treat02-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a justiça
é muito lenta e os criminosos ficam soltos até o julgamento"**.

@@

* You will be shown a news article. Please read it carefully. After you read
  the article, we will ask you one question about it.[^vig01-exp02]
